Inducing Semantic Frames on a Very Large Corpus of Syntactic-Ngrams

### 2015-01-05 Paper/paperN.tex
final version

\begin{abstract}
We explore two models for the unsupervised induction of semantic frames on a 
very large corpus of verb-subject-object triples. Our work is bade on previous 
research that has had success on smaller, more focused data sets. By encouraging 
verb sparsity with a Dirichlet prior, one can improve on the results of our 
baseline model. We give some analysis of the frames that the improved model 
discovers, and suggest areas for further development.

  Information in text can be represented by frames. A frame consists of a predicate with arguments. In this report we discuss two probabilistic latent variable models which are trained by resp. expectation maximization and Gibbs sampling, to induce frames from syntactic tri-grams of the form: verb, subject, object. The latent variable is the frame a verb belongs to. In practice, the models cluster verbs based on the occurrences of arguments they are encountered with. The frames (or clusters of verbs) are measured by correctness (relative to each other) and correctness (relative to FrameNet clusters). The more suffisticated second model performs well and finds clusters of verbs which have a distinct general meaning, the frames induced with the simple model 0 are not as well defined as thos of model 1. 
\end{abstract}

@@

Semantic content of a sentence can be represented in a frame. The main verb phrase of the sentence is the head of the frame and can be interpreted as predicate, representing the core of the episode described by the sentence. The trick is to find out what the arguments of the predicate are. In linguistics, argument selection is called case grammar \citep{dowty1991} and arguments (or roles) have specific names as \textit{agent - patient}. In computational systems names are not important and can be assigned when frame induction is complete.

@@

%\section{Related Work}
We based our work on two papers: \textit{`Inducing a semantically annotated lexicon via em-based clustering.'} by \citeauthor{rooth1999} and \textit{`Learning frames from text with an unsupervised latent variable model.'} by \citeauthor{oconnor2013}. In the first article clustering is realized by applying expectation maximalization on a simple cooccurance model. This model produces frames in the form of two distributions, one for the predicate, a distribution over verbs, and one for the arguments, a nouns. Following O'Connor, we modify this model slightly to accomodate verb-subject-object triples. A triple is assigned to the frame for which is probability of occuring is highest. We will use this model as our baseline (as in the O'Connor paper). The second paper covers two additional models, of which we implemented one. This model differs from the baseline model by introducing the assumption that a document generates sentences frome a restricted set of frames. This assumption is justified (as in latent Dirichelt allocation) \cite{blei2003} by the intuition that documents generally have a limited number of topics and thus use a proper subset of the frames. In our data the tuples are not annotated with documents so we used this assumption in such a way that each verb covers one document.

@@

It is in the nature of documents that they have some degree of internal semantic
coherence. Put more plainly, a document is usually \emph{about} something. We
know from topic modeling that LDA with Gibbs sampling can be very successful at
classifying documents according to some semantic features \citep{blei2003}.
Model 1 leverages this fact by making the assumption that a given document is 
likely to have a sparse distribution of frames. As in LDA, the document-level
sparsity is enforced by a Dirichlet prior on frame distributions (see figure
\ref{gen1} for details). This assumption breaks the independence between tuples 
that is maintained by Model 0 is dropped in Model 1.

@@

\begin{align*}
 P(\mathbf{w},\mathbf{f}|\alpha,\beta)
&=\int_\phi\int_\theta P(\mathbf{w},\mathbf{f}|\theta,\phi,\alpha,\beta)\\
&=     \int_\phi P(\mathbf{w}|\mathbf{f},\phi )\,P(\phi|\beta)
\times \int_\theta P(\mathbf{f}|\theta)\,P(\theta|\alpha)\\
\intertext{These integrals are both Dirichlet multinomials:}
&=\frac{DCM_i^{frames}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{frames}(\mathbf{f}_{-ij})}\times
\prod_a^{\{v,s,o\}}\frac{DCM_i^{vocab^a}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{vocab^a}(\mathbf{f}_{-ij})}
\end{align*}

@@

The data used in this project (see section \ref{data}) doesn't contain any document-level
information. As a result, we adapt model 1 by grouping VSOs with the same verb into ``documents''.
This move leverages the Dirichlet prior $\alpha$ to enforce sparsity of frames 
across verbs, but of course cannot capture the information given by the kind of
document-level semantic sparsity seen in LDA.\footnote{When we use verbs as
documents (i.e, $D = V^v$), since $w_{ij}^v$ is the same for all verbs in document $d_i$ 
(and $d_i$ is the only document with that verb), it follows that the number of VSO's in document 
$d_i$ belonging to frame $f$ is the same as the number of VSO's with $w_{ij}^v$ as their verb assigned to 
frame $f$; that is, $\tilde c(d_i, f) = \tilde c(f, w_{ij}^v)$.}

@@

A syntactic-ngram is a $k$-word rooted subtree of some sentence.
Google syntactic-ngrams is a corpus of ngrams extracted from 3.5 million English language 
books and labled with Penn Treebank part-of-speech tags Stanford-basic dependency 
labels \citep{ngrams2013}.
Each ngram also includes frequency counts by year.

This project uses the verbargs subeset (130M items) of Google syntactic-ngrams.
An ngram in the verbargs dataset consist of a verb and all of its 
immediate arguments.
We considered only the VSO's in this dataset, that is; we pruned verbargs to
those ngrams containing a subject (nsubj) and object (dobj) whose immediate parent
is the ngram's root verb (1.6M items, 96M by count). 
We further trimmed the data to consider only the \%25 most common verbs 
(800K items, 78M by count).

@@

The $\alpha$ and $\beta$ hyperparameters in model 1 are known  to be hard to determine \citep{oconnor2013}, and therefore need to be experimentally optimized for our specific problem. We used an expensive, but easy to implement grid search to determine the best values for $\alpha$ and $\beta$. The grid search builds, for each posible parameter set, a model which is evaluated on the cross validation set. The highest scoring model is tested with the test set. We ran grid search for model 0 on $50$ - $200$ frames, with steps of $50$, and $\alpha$ of $0.1, 0.5, 1, 1.5$, and $2$. The grid search for Model 1 was the the same, additionally testing $\beta$ across the same values as $\alpha$
\section{Results}
\label{results}

\subsection{Frame Coherency}
Verb clusters are based on the arguments with which they occurring in the data. So to test how well the models perform on clustering, the probability that the model generates a triple from the testset, should be greater than this triple with a random verb in place of the real one. This metric is the same as the one used by \citeauthor{rooth1999}. More formally: for a datapoint $(v,s,o)$ and a tuple $(v^r,s,o)$ where $v^r$ is a random chosen verb: $P(v\mid s,o) \geq P(v^r\mid s,o)$.
The results are percentages of data points which satisfy this condition. Note though that if the random verb-triple is assigned the same frame as its `parent' real triple, this metric loses its meaning, but since the number of frames is high this probability is low. Furthermore, one can assume that the occurrences of same frame assignment cancel each other out, i.e. the random triple has higher probability than the real one as many times as vice versa.

@@

%Frame correctness: for the top 25 most probable verbs per frame $TV$ and framenet classes of verbs $FN$: \[\frac{2|TV\cap FN|}{|TV|+|FN|}\]
\section{Results}
These results are experimental, i.e. the parameters are guessed and therefor probably not optimal. (Since gridsearchis more expensive than planned for the actual results are still in progress.) In this experiment the frames were put on $200$, $\alpha$ and $\beta$ both on $0.5$.\\
\begin{table}[h]
\begin{tabular}{|l||c|c|}
  \hline
  &dicesim&coherencey\\
  \hline
  \hline
  model 0& 0.28&19\%  \\
  \hline
  model 1&0.29 &71\%  \\
  \hline
\end{tabular}
\caption{results}
\label{results}
\end{table}

@@

\section{Discussion}
As table \ref{results} show, coherency  improved significantly, whereas the frame correctness did not so much. But the frames induced by model 0 do not cover all verbs, actually only 3 verbs are asigned to a frame. It is not clear why this happens, but this are still only experimental results.  Since there are 200 frames it is obvious that the coherency is so low. Other experiments have shown that it is possible to cluster with model 0 though.

The frames in table \ref{exframes} are frames induced by model 1 and are meaningfull in the sense that they consist of clusters of verbs that seem to have a similar meaning. Nontheless, in \textit{harming anotherly} there are verbs which would be better left out. They are assigned this frame though because they occurred in company of the same arguments often. The affection frame is remarkable since these are exactly opposites. This happens in some form in more frames though and can be explained by the fact that verbs with opposite meanings are used with the same arguments, this can be seen as a weakness of the model.

Furthermore auxilary verbs get mostly assigned a complete frame, i.e. they are the only verb assigned to a frame. Other verbs like \textit{(say, get, find, bring, make, see, ask, take, have, mean, use, like, hear, understand, afford, help, call, know, give, do, tell, find)} also get assigned a complete frame. This is most likely caused by the fact that these verbs can have a wide range of arguments, and are in the sense of meaning-of-a-verb by nouns it is encountered with, sort of meaningless. This is again a weakness of the model, but could possibly be diminished with more data.

\section{Future Work}
One big weakness of this project over the results of \citet{oconnor2013} is that 
we could not used document-level context to infer semantic information.
A possible direction of future research is to create models that take better 
advantage of the syntax-level context provided by syntactic ngrams. 
One easy way to do this would be to modify model 1 so that the argument 
vocabulary ranges over subject/object pairs instead of individual nouns.
or object.
More advanced models might take better advantage of the various kinds of arguments
in the verbargs dataset and their relation to one another.
For example, you could be glean some semantic information about the relationship between
verbs and prepositional objects by looking at the preposition that mediates them: 
That you ''swim \emph{in} a lake'' versus ``ice skate \emph{on} a lake'' 
indicates that the relationship between swimming and lakes is different from the one between
ice skating and lakes.

\section{Conclusion}

\section*{Who did what}
We all tried to work as much as possible and neccesary on the project. Bill had a leading role in building the models. Model 1 is completely build and implemented by Bill. Jacob implemented frame coherency and Eli frame correctness. Everybody ran some experiments and Jacob is still running grid search (made a big estimation error in cost of gridsearch). Any section is written by the one who implemented this particular part. The rest of the sections are composed of bits of everybody.

>>>>

\begin{abstract}
  Information in text can be represented by frames. A frame consists of a predicate with arguments. In this report we discuss two probabilistic latent variable models which are trained by resp. expectation maximization and Gibbs sampling, to induce frames from syntactic tri-grams of the form: verb, subject, object. The latent variable is the frame a verb belongs to. In practice, the models cluster verbs based on the occurrences of arguments they are encountered with. The frames (or clusters of verbs) are measured by correctness (relative to each other) and correctness (relative to FrameNet clusters). The more sophisticated second model performs well and finds clusters of verbs which have a distinct general meaning, the frames induced with the simple model 0 are not as well defined as those of model 1. 
\end{abstract}

@@

Semantic content of a sentence can be represented in a frame. The main verbphrase of the sentence is the head of the frame and can be interpreted as predicate, representing the core of the episode described by the sentence. The trick is to find out what the arguments of the predicate are. In linguistics, argument selection is called case grammar \citep{dowty1991} and arguments (or roles) have specific names as \textit{agent - patient}. In computational systems names are not important and can be assigned when frame induction is complete.

@@

%\section{Related Work}
We based our work on two papers: \textit{`Inducing a semantically annotated lexicon via em-based clustering.'} by \citeauthor{rooth1999} and \textit{`Learning frames from text with an unsupervised latent variable model.'} by \citeauthor{oconnor2013}. In the first article clustering is realized by applying expectation maximalization on a simple co-occurance model. This model produces frames in the form of two distributions, one for the predicate, a distribution over verbs, and one for the arguments, a nouns. Following O'Connor, we modify this model slightly to accommodate verb-subject-object triples. A triple is assigned to the frame for which is probability of occurring is highest. We will use this model as our baseline (as in the O'Connor paper). The second paper covers two additional models, of which we implemented one. This model differs from the baseline model by introducing the assumption that a document generates sentences from a restricted set of frames. This assumption is justified (as in latent Dirichlet allocation) \cite{blei2003} by the intuition that documents generally have a limited number of topics and thus use a proper subset of the frames. In our data the tuples are not annotated with documents so we used this assumption in such a way that each verb covers one document.

@@

It is in the nature of documents that they have some degree of internal semantic
coherence. Put more plainly, a document is usually \emph{about} something. We
know from topic modeling that LDA with Gibbs sampling can be very successful at
classifying documents according to some semantic features \citep{blei2003}.
Model 1 leverages this fact by making the assumption that a given document is 
likely to have a sparse distribution of frames. As in LDA, the document-level
sparsity is enforced by a dirichlet prior on frame distributions (see figure
\ref{gen1} for details). This assumption breaks the independence between tuples 
that is maintained by Model 0 is dropped in Model 1.

@@

\begin{align*}
 P(\mathbf{w},\mathbf{f}|\alpha,\beta)
&=\int_\phi\int_\theta P(\mathbf{w},\mathbf{f}|\theta,\phi,\alpha,\beta)\\
&=     \int_\phi P(\mathbf{w}|\mathbf{f},\phi )\,P(\phi|\beta)
\times \int_\theta P(\mathbf{f}|\theta)\,P(\theta|\alpha)\\
\intertext{These integrals are both dirichlet multinomials:}
&=\frac{DCM_i^{frames}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{frames}(\mathbf{f}_{-ij})}\times
\prod_a^{\{v,s,o\}}\frac{DCM_i^{vocab^a}(f_{ij}=f, \mathbf{f}_{-ij})}{DCM_i^{vocab^a}(\mathbf{f}_{-ij})}
\end{align*}

@@

The data used in this project (see section \ref{data}) doesn't contain any document-level
information. As a result, we adapt model 1 by grouping VSOs with the same verb into ``documents''.
This move leverages the dirichlet prior $\alpha$ to enforce sparcity of frames 
across verbs, but of course cannot capture the information given by the kind of
document-level semantic sparsity seen in LDA.\footnote{When we use verbs as
documents (i.e, $D = V^v$), since $w_{ij}^v$ is the same for all verbs in document $d_i$ 
(and $d_i$ is the only document with that verb), it follows that the number of VSO's in document 
$d_i$ belonging to frame $f$ is the same as the number of VSO's with $w_{ij}^v$ as their verb assigned to 
frame $f$; that is, $\tilde c(d_i, f) = \tilde c(f, w_{ij}^v)$.}

@@

A syntactic-ngram is a $k$-word rooted subtree of some sentence.
Google syntactic-ngrams is a corpus of ngrams extracted from 3.5 million English language 
books and labeled with Penn Treebank part-of-speech tags Stanford-basic dependency 
labels \citep{ngrams2013}.
Each ngram also includes frequency counts by year.

This project uses the verbargs subset (130M items) of Google syntactic-ngrams.
An ngram in the verbargs dataset consist of a verb and all of its 
immediate arguments.
We considered only the VSO's in this dataset, that is; we pruned verbargs to
those ngrams containing a subject (nsubj) and object (dobj) whose immediate parent
is the ngram's root verb (1.6M items, 96M by count). 
We further trimmed the data to consider only the \%25 most common verbs 
(800K items, 78M by count).

@@

The $\alpha$ and $\beta$ hyper-parameters in model 1 are known  to be hard to determine \citep{oconnor2013}, and therefore need to be experimentally optimized for our specific problem. We used an expensive, but easy to implement grid search to determine the best values for $\alpha$ and $\beta$. The grid search builds, for each possible parameter set, a model which is evaluated on the cross validation set. The highest scoring model is tested with the test set. We ran grid search for model 0 on $50$ - $200$ frames, with steps of $50$, and $\alpha$ of $0.1, 0.5, 1, 1.5$, and $2$. The grid search for Model 1 was the the same, additionally testing $\beta$ across the same values as $\alpha$
\section{Results}
\label{results}

\subsection{Frame Coherency}
Verb clusters are based on the arguments with which they occurring in the data. So to test how well the models perform on clustering, the probability that the model generates a triple from the test-set, should be greater than this triple with a random verb in place of the real one. This metric is the same as the one used by \citeauthor{rooth1999}. More formally: for a data-point $(v,s,o)$ and a tuple $(v^r,s,o)$ where $v^r$ is a random chosen verb: $P(v\mid s,o) \geq P(v^r\mid s,o)$.
The results are percentages of data-points which satisfy this condition. Note though that if the random verb-triple is assigned the same frame as its `parent' real triple, this metric loses its meaning, but since the number of frames is high this probability is low. Furthermore, one can assume that the occurrences of same frame assignment cancel each other out, i.e. the random triple has higher probability than the real one as many times as vice versa.

@@

%Frame correctness: for the top 25 most probable verbs per frame $TV$ and framenet classes of verbs $FN$: \[\frac{2|TV\cap FN|}{|TV|+|FN|}\]
\section{Results}
These results are experimental, i.e. the parameters are guessed and therefor probably not optimal. (Since grid-search is more expensive than planned for the actual results are still in progress.) In this experiment the frames were put on $200$, $\alpha$ and $\beta$ both on $0.5$.\\
\begin{table}[h]
\begin{tabular}{|l||c|c|}
  \hline
  &dicesim&coherency\\
  \hline
  \hline
  model 0& 0.28&19\%  \\
  \hline
  model 1&0.29 &71\%  \\
  \hline
\end{tabular}
\caption{results}
\label{results}
\end{table}

@@

\section{Discussion}
As table \ref{results} show, coherency  improved significantly, whereas the frame correctness did not so much. But the frames induced by model 0 do not cover all verbs, actually only 3 verbs are assigned to a frame. It is not clear why this happens, but this are still only experimental results.  Since there are 200 frames it is obvious that the coherency is so low. Other experiments have shown that it is possible to cluster with model 0 though.

The frames in table \ref{exframes} are frames induced by model 1 and are meaningful in the sense that they consist of clusters of verbs that seem to have a similar meaning. Nonetheless, in \textit{harming anotherly} there are verbs which would be better left out. They are assigned this frame though because they occurred in company of the same arguments often. The affection frame is remarkable since these are exactly opposites. This happens in some form in more frames though and can be explained by the fact that verbs with opposite meanings are used with the same arguments, this can be seen as a weakness of the model.

Furthermore auxiliary verbs get mostly assigned a complete frame, i.e. they are the only verb assigned to a frame. Other verbs like \textit{(say, get, find, bring, make, see, ask, take, have, mean, use, like, hear, understand, afford, help, call, know, give, do, tell, find)} also get assigned a complete frame. This is most likely caused by the fact that these verbs can have a wide range of arguments, and are in the sense of meaning-of-a-verb by nouns it is encountered with, sort of meaningless. This is again a weakness of the model, but could possibly be diminished with more data.

\section{Future Work}
One big weakness of this project over the results of \citet{oconnor2013} is that 
we could not used document-level context to infer semantic information.
A possible direction of future research is to create models that take better 
advantage of the syntax-level context provided by syntactic ngrams. 
One easy way to do this would be to modify model 1 so that the argument 
vocabulary ranges over subject/object pairs instead of individual nouns.
or object.
More advanced models might take better advantage of the various kinds of arguments
in the verbargs dataset and their relation to one another.
For example, you could glean some semantic information about the relationship between
verbs and prepositional objects by looking at the preposition that mediates them: 
That you ''swim \emph{in} a lake'' versus ``ice skate \emph{on} a lake'' 
indicates that the relationship between swimming and lakes is different from the one between
ice skating and lakes.

\section{Conclusion}
Model 1 did find clusters of verbs of which the meaning can be covered by 1 predicate. Although the induced frames have often vague or broad meaning the approach works well if data is abundant. The improvement of model 1 on model 0 has a significant effect and implicates even better results can be acquired by further improvements. The kind of clustering done by these models, with opposites assigned to the same frame, can be usefull for automatic stock trading systems if implemented with sentiment analysis to split on opposites. The frames are too vague to use for summarizations though.

\section*{Who did what}
We all tried to work as much as possible and necessary on the project. Bill had a leading role in building the models. Model 1 is completely build and implemented by Bill. Jacob implemented frame coherency and Eli frame correctness. Everybody ran some experiments and Jacob is still running grid search (made a big estimation error in cost of grid-search). Any section is written by the one who implemented this particular part. The rest of the sections are composed of bits of everybody.
